Synthesizing Symmetric Lenses

### 2018-05-29 papers/SymmetricLenses/main.tex
add in related work and stuff

% begin related-work
\section{Related Work}
\label{sec:related}
There is related work in a number of directions.  There is related work in
synthesizing synchronization functions, related work on symmetric lenses,
related work on string transformation synthesis, and related work on
type-directed synthesis.

\subsection{Symmetric Lenses}
We have reformulated the problem addressed by symmetric lenses with simple
symmetric lenses.

\section{Simple Symmetric Lenses}
\label{sec:ssl2}
Intuitively, simple symmetric lenses are much like symmetric lenses, but they
avoid complements by keeping all information required to stitch edits to a data
source in the data source itself. Instead of having puts operate on a
complement, they operate on both data sources, where the alterations of the
first argument update the second. Furthermore, on initial creations of the
files, instead of using an $init$ complement, two additional functions,
$\CreateR$ and $\CreateL$ are used.

More formally, a simple symmetric lens between $X$ and $Y$ comprise the
following four functions:
\begin{enumerate}
\item $\CreateR \OfType X \rightarrow Y$
\item $\CreateL \OfType Y \rightarrow X$
\item $\PutR \OfType X \rightarrow Y \rightarrow Y$
\item $\PutL \OfType Y \rightarrow X \rightarrow X$
\end{enumerate}
These functions are guaranteed to satisfy the following ``round-tripping'' laws.
\begin{gather}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{gather}

\section{Symmetric Lenses}
\label{sec:symmetric-lenses}

Classically, these types of synchronization tasks are represented as symmetric
lenses. A symmetric lens $\Lens$ between $X$ and $Y$ consists of 4 components: a
complement $C$, a designated element $init \in C$, and two functions for
propagating changes in each format to the other:
\begin{enumerate}
\item $\PutRSym \OfType X \times C \rightarrow Y \times C$
\item $\PutLSym \OfType Y \times C \rightarrow X \times C$
\end{enumerate}
In this formulation, the components of each format that aren't shared with the
other format are stored in the complement. When one format is edited, the \PutR
or \PutL function stitches together that edited data with the data for the other
side, stored in the complement.  The designated element, $init$, is used as the
initial value of $C$, and specifies default behavior when there isn't existing
data to stitch with.  In our example, the complement would consist of two lists
of commands, where one list contains the Linux commands and one list contains
the Windows commands.

These components only make sense for synchronization when they satisfy
round-tripping laws.  This guarantees that if an identity edit is propagated, no
data gets changed.
\begin{equation}
  \tag{\PutRL}
  \frac{\PutRSymOf{(x,c)} = (y,c')}{\PutLSymOf{(y,c''''')} = (x,c')}
\end{equation}
\begin{equation}
  \tag{\PutLR}
  \frac{\PutLSymOf{(y,c)} = (x,c''')}{\PutRSymOf{(x,c')} = (y,c')}
\end{equation}

@@

\subsection{Difficulties Posed}

@@

Using symmetric lenses for synchronization forces us to maintain not just two
files that we aim to sycnrhonize, but three.  This is unintuitive, and creates
many problems in practice.  For example, consider the situation where I am
synchronizing a crontab and job manager file.  I am synchronizing across
computers, so I must either maintain a copy of the complement on each computer,
and keep them synchronized, or figure out a computer to hold the complement, and
make sure all edits go through that complement.  Furthermore, if I want to
change the synchronization to a different computer, I don't need to merely make
sure that my files are moved to the computer, I also have to make sure that I
correctly move the complement, and hook up that complement to be used for that
synchronization task.  This is a huge divergence from other synchronization
utilities, where all that needs to be done is maintain the files.  Using those
for synchronization, edits can be sent over the wire, and the appropriate
function can stitch together those edits, no additional state needed.

@@

A \emph{symmetric lens} between $X$ and $Y$ is an element of the type
$\exists C.\{ \mathit{missing} :
C,\,get : X \times C \to Y \times C,\,put : Y \times C \to X \times C\}$
satisfying the following laws:
\begin{equation}
  \tag{\PutRL}
  \frac{\PutR(x,c) = (y,c')}{\PutL(y,c') = (x,c')}
\end{equation}
\begin{equation}
  \tag{\PutLR}
  \frac{\PutL(y,c) = (x,c')}{\PutR(x,c') = (y,c')}
\end{equation}

Given a lens $l \in X \leftrightarrow Y$, a \emph{put object} for $l$ is a
member of $X + Y$. Consider the function $apply$, which, given a lens and an
element of that lens's complement, is a function from sequences of put objects
to sequences of put objects. If $apply(\Lens,c,es) = es'$, then that means, starting
with complement $c$, after edit $es_i$, the lens $\Lens$ synchronizes the other
data to $es'_i$.
\[
  \inferrule
  {
  }
  {
    apply(\Lens,c,[]) = []
  }
\]
\[
  \inferrule
  {
    \Lens.putr(x,c) = (y,c')\\
    apply(\Lens,c',es) = es'
  }
  {
    apply(\Lens,c,(\InLOf{x})::es) = (\InROf{y})::es'
  }
\]
\[
  \inferrule
  {
    \Lens.putl(y,c) = (x,c')\\
    apply(\Lens,c',es) = es'
  }
  {
    apply(\Lens,c,(\InROf{y})::es) = (\InLOf{x})::es'
  }
\]

Two simple symmetric lenses are equivalent if the four functions they comprise
are equivalent as functions. However, to formalize how these lenses work as
synchronization tools, we define an $apply$ function on them, much like we
define an $apply$ on symmetric lenses.

The function $apply$, given a symmetric lens and an optional pair of elements of
the source and target, is a function from sequences of put objects to sequences
of put objects.  If $apply(\Lens,\None,es,es')$, then that means, starting with
no prior data, after edit $es_i$, the lens \Lens synchronizes the other data to
$es_i'$.  If $apply(\Lens,\SomeOf{(x,y)},es,es')$, then that means, starting with
data $x$ and $y$ on the left and right, respectively, after edit $es_i$, the
lens \Lens synchronizes the other data to $es_i'$.

@@

Furthermore, equivalence of the $apply$ function corresponds exactly with
equivalence of lenses.

\begin{theorem}
  $\Lens_1 \equiv \Lens_2$ if, and only if, for all sequences $es$,
  $apply(\Lens_1,\None,es) = apply(\Lens_2,\None,es)$.
\end{theorem}

@@

\begin{theorem}
  Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,
  where
  \begin{enumerate}
  \item $\Lens.creater \App x = \Lens.get \App x$
  \item $\Lens.createl \App y = \Lens.create \App y$
  \item $\Lens.putr \App x \App y = \Lens.get \App x$
  \item $\Lens.putl \App y \App x = \Lens.put \App y \App x$
  \end{enumerate}
\end{theorem}

>>>>

% begin related-work
\section{Related Work}
\label{sec:related}
There is related work in a number of directions.  There is related work in
symmetric lenses, related work on synthesizing synchronization functions,
related work on string transformation synthesis, and related work on
type-directed synthesis.

\subsection{Symmetric Lenses}
\paragraph*{Formalization} We have reformulated the problem addressed by
symmetric lenses with simple symmetric lenses. Classically, these types of
synchronization tasks are represented as symmetric lenses. A symmetric lens
$\Lens$ between $X$ and $Y$ consists of 4 components: a complement $C$, a
designated element $init \in C$, and two functions for propagating changes in
each format to the other:
\begin{enumerate}
\item $\PutRSym \OfType X \times C \rightarrow Y \times C$
\item $\PutLSym \OfType Y \times C \rightarrow X \times C$
\end{enumerate}
In this formulation, the components of each format that aren't shared with the
other format are stored in the complement. When one format is edited, the \PutR
or \PutL function stitches together that edited data with the data for the other
side, stored in the complement.  The designated element, $init$, is used as the
initial value of $C$, and specifies default behavior when there isn't existing
data to stitch with.  In our example, the complement would consist of two lists
of commands, where one list contains the Linux commands and one list contains
the Windows commands.

These components only make sense for synchronization when they satisfy
round-tripping laws.  This guarantees that if an identity edit is propagated, no
data gets changed.
\begin{equation}
  \tag{\PutRL}
  \frac{\PutRSymOf{(x,c)} = (y,c')}{\PutLSymOf{(y,c')} = (x,c')}
\end{equation}
\begin{equation}
  \tag{\PutLR}
  \frac{\PutLSymOf{(y,c)} = (x,c')}{\PutRSymOf{(x,c')} = (y,c')}
\end{equation}

@@

@@

\paragraph*{Issues with Symmetric Lenses} Using symmetric lenses for
synchronization forces us to maintain not just two files that we aim to
sycnrhonize, but three. This is unintuitive, and creates many problems in
practice. For example, consider the situation where I am synchronizing a crontab
and job manager file. I am synchronizing across computers, so I must either
maintain a copy of the complement on each computer, and keep them synchronized,
or figure out a computer to hold the complement, and make sure all edits go
through that complement. Furthermore, if I want to change the synchronization to
a different computer, I don't need to merely make sure that my files are moved
to the computer, I also have to make sure that I correctly move the complement,
and hook up that complement to be used for that synchronization task. This is a
huge divergence from other synchronization utilities, where all that needs to be
done is maintain the files. Using those for synchronization, edits can be sent
over the wire, and the appropriate function can stitch together those edits, no
additional state needed.

@@

\paragraph*{Relation with Simple Symmetric Lenses} To formalize how simple
symmetric lenses lenses work as synchronization tools, we define an $apply$
function on them, much like we define an $apply$ on symmetric lenses. The
function $apply$, given a symmetric lens and an optional pair of elements of the
source and target, is a function from sequences of put objects to sequences of
put objects. If $apply(\Lens,\None,es,es')$, then that means, starting with no
prior data, after edit $es_i$, the lens \Lens synchronizes the other data to
$es_i'$. If $apply(\Lens,\SomeOf{(x,y)},es,es')$, then that means, starting with
data $x$ and $y$ on the left and right, respectively, after edit $es_i$, the
lens \Lens synchronizes the other data to $es_i'$.

@@

@@

\begin{theorem}
  Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,
  where
  \begin{center}
    \begin{tabular}{rcl}
      $\Lens.\CreateROf{x}$ & $=$ & $\Lens.get \App x$\\
      $\Lens.\CreateLOf{y}$ & $=$ & $\Lens.create \App y$\\
      $\Lens.\PutROf{x}{y}$ & $=$ & $\Lens.get \App x$\\
      $\Lens.\PutLOf{y}{x}$ & $=$ & $\Lens.put \App y \App x$
    \end{tabular}
  \end{center}
\end{theorem}

\subsection{Synthesizing Synchronization Functions}

\subsection{String Transformation Synthesis}

\subsection{Type-Directed Synthesis}
